A Constraint Solving Approach to Parikh Images of Regular Languages

### 2022-07-08 main.tex
Experiments

We evaluate the performance of \Catra{} on~\NrBenchmarks{} instances of Parikh
automata intersection problems generated by the \OstrichPlus{} string constraint
solver when solving the PyEx benchmarks~\cite{pyex}. After generating an
initial~\InitialNrBenchmarks{}, we remove misgenerated instances that did not
parse, as well as~\NrTrivial{} instances solved in under five seconds by
baseline. The benchmarks are run on commit~\texttt{\commit}.

@@

In \cref{fig:solve-division} we show how many of the~\NrBenchmarks{} instances
the respective back-ends could solve and with which status. A full summary of
their outcomes is also available in \cref{tab:solve-status}. We see that
\Calculus{} generally outperforms \Nuxmv{} on unsatisfiable instances, while
being worse at satisfiable instances. Both \Nuxmv{} and \Calculus{} strictly
outperform baseline on every kind of instance, but most of all on satisfiable
instances. \Cref{tab:solve-status} also shows a small number of bugs in
\Calculus{}: 27~satisfiable instances are misreported as \Unsat{}, and 20
additional instances fail with an error. Both problems are due to
non-deterministic bugs in the clause learning. Note that \Calculus{} is sound:
it never misreports an unsatisfiable instance as \Sat{}, and all
\Fudge{\numprint{6218}} satisfying assignments have been verified by \Nuxmv{}.
Baseline performs worse on satisfiable instances because it executes a heuristic
meant to detect unsatisfiability early at a heavy penalty to satisfiable
instances.

@@

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{graphs/\commit-by-solver.pdf}
    \caption{The division of statuses per backend.}
    \label{fig:solve-division}
    \Description[A bar chart showing three bars, one per backend, illustrating how many instances they could solve]{The bars are divided by satisfiable and unsatisfiable instances. Baseline could seemingly only solve unsatisfiable instances, lazy could solve a few satisfiable and mostly unsatisfiable, and nuxmv could solve about twice as many unsatisfiable as satisfiable instances, and slightly more in total than lazy.}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{graphs/\commit-time-boxplot.pdf}
    \caption{The distribution of runtimes for solved instances per backend. Note that the number of instances solved differs between backends.}
    \label{fig:runtime-boxplot}
    \Description[A box plot showing the distribution of runtime over the three backends]{The middle box plot shows a tiny box centered around 0 seconds for the lazy backend, the rightmost box shows a bigger box between five seconds and 30 seconds for nuxmv, and the leftmost box shows a smaller box between 20 and 30 seconds for baseline. The whiskers of nuxmv span between 0 and the timeout, 60 seconds, while they are much tighter for lazy. On the other hand, lazy has  a lot of outliers.}
  \end{subfigure}
  \vfill
  \begin{subfigure}[b]{0.75\textwidth}
    \includegraphics[width=\textwidth]{graphs/\commit-cactus.pdf}
    \caption{The number of instances solved as the time budget increases, simulated from one \RuntimeTimeout-timeout run.}
    \label{fig:cactus}
    \Description[A cactus plot comparing the performance of nuxmv, lazy and baseline]{The line for baseline is strictly lower than the two others, and does not head upwards from zero until after 10 seconds of timeout, while the number of instances for nuxmv increases sharply until ca 10 seconds and then linearly after that. Lazy solves much more instances than nuxmv on shorter runtimes, but is overtaken by just before the 60-second timeout used in these experiments.}
  \end{subfigure}
\end{figure}

@@

A cactus plot showing the number of instances solved within a given timeout for
each backend can be seen in \cref{fig:cactus}. Here, we see clearly that
\Calculus{} outperforms \Nuxmv{} for shorter timeouts, and that \Nuxmv{} is
only able to catch up once we increase the timeout to far beyond what would be
normal for a string solver.

>>>>

We evaluate the performance of \Catra{} on~\NrBenchmarks{} instances of Parikh
automata intersection problems generated by the \OstrichPlus{} string constraint
solver when solving the PyEx benchmarks~\cite{pyex}. After generating an
initial~\InitialNrBenchmarks{},we remove~\NrTrivial{} instances solved in under
five seconds by baseline. We also attempted to benchmark instances generated by \Ostrich{} solving the kaluza-len (\numprint{38227} instances) and pyex-len (\numprint{791} instances) benchmark suites, but had to discard them since they were all trivial. For instance, on a laptop, \Calculus{} could solve every instance from kaluza-len in under five seconds, mean about 0.1s. The benchmarks are run on commit~\texttt{\commit}.

@@

In \cref{fig:solve-division} we show how many of the~\NrBenchmarks{} instances left after excluding the trivial instances, among them the \numprint{38227} and \numprint{791} instances generated from kaluza-len and pyex-len respectively,
the respective back-ends could solve and with which status. A full summary of
their outcomes is also available in \cref{tab:solve-status}. Note that we lack
ground truth for all instances. We see that \Calculus{} generally outperforms
\Nuxmv{} on determining unsatisfiability, while being worse at satisfiable
instances. Both \Nuxmv{} and \Calculus{} strictly outperform baseline on every
kind of instance, but most of all on satisfiable instances, and \Calculus{}
outperforms \Nuxmv{}. \Nuxmv{} and \Calculus{} disagree on 38 instances, where
\Nuxmv{} returns \Sat{} and \Calculus{} returns \Unsat{}. We believe this to be
caused by a non-deterministically triggered bug in the clause learning of
\Calculus{}, but have not had time to further investigate it. We have however
validated all \numprint{3908} satisfying assignments from \Calculus{} with
\Nuxmv{}. \Cref{tab:solve-status} also shows a small number of crashing bugs in
our implementation.

Baseline performs worse on satisfiable instances because it executes a heuristic
meant to detect unsatisfiability early at a heavy penalty to satisfiable
instances. The heuristic is enabled since it is unrealistic to assume that we
know in advance whether an instance is satisfiable or not, but we do know that
unsatisfiable instances are more common.

@@

\begin{figure}[t]
  \centering
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{graphs/\commit-by-solver.pdf}
    \caption{The division of statuses per backend.}
    \label{fig:solve-division}
    \Description[A bar chart showing three bars, one per backend, illustrating how many instances they could solve]{The bars are divided by satisfiable and unsatisfiable instances. Baseline could seemingly only solve unsatisfiable instances, lazy could solve a few satisfiable and mostly unsatisfiable, and nuxmv could solve about twice as many unsatisfiable as satisfiable instances, and slightly less in total than lazy.}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.49\textwidth}
    \includegraphics[width=\textwidth]{graphs/\commit-time-boxplot.pdf}
    \caption{The distribution of runtimes for solved instances per backend. Note that the number of instances solved differs between backends.}
    \label{fig:runtime-boxplot}
    \Description[A box plot showing the distribution of runtime over the three backends]{The middle box plot shows a tiny box centered around 0 seconds for the lazy backend, the rightmost box shows a bigger box between five seconds and 30 seconds for nuxmv, and the leftmost box shows a smaller box between 20 and 30 seconds for baseline. The whiskers of nuxmv span between 0 and the timeout, 60 seconds, while they are much tighter for lazy. On the other hand, lazy has  a lot of outliers.}
  \end{subfigure}
  \vfill
  \begin{subfigure}[b]{0.75\textwidth}
    \includegraphics[width=\textwidth]{graphs/\commit-cactus.pdf}
    \caption{The number of instances solved as the time budget increases, simulated from one \RuntimeTimeout-timeout run.}
    \label{fig:cactus}
    \Description[A cactus plot comparing the performance of nuxmv, lazy and baseline]{The line for baseline is strictly lower than the two others, and does not head upwards from zero until after 10 seconds of timeout, while the number of instances for nuxmv increases sharply until ca 10 seconds and then linearly after that. Lazy solves much more instances than nuxmv.}
  \end{subfigure}
\end{figure}

@@

A cactus plot showing the number of instances solved within a given timeout for
each backend can be seen in \cref{fig:cactus}. Here, we see clearly that
\Calculus{} outperforms \Nuxmv{} in general, but that \Nuxmv{} might have better
long-term scaling properties.
